Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology

### 2015-04-29 article-00-headers.tex
Wrote sth about ifcRDF

\spnewtheorem{ontologyRule}{Ontology Rule}{\bfseries}{\rmfamily}

\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}

@@

% ifcOWL layer names
\newcommand{\lite}{\textbf{Lite}}
\newcommand{\standard}{\textbf{Standard}}
\newcommand{\advanced}{\textbf{Advanced}}

\newcommand{\ifclite}{\textbf{ifcOWL-Lite}}
\newcommand{\ifcstandard}{\textbf{ifcOWL-Standard}}
\newcommand{\ifcadvanced}{\textbf{ifcOWL-Advanced}}

@@

%\vspace{1cm}
%
\title{Publishing Datasets about Building Designs \\Based on a Multilayer Building Ontology}
%
\titlerunning{Building ontology}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Nam Vu Hoang \and Seppo T\"orm\"a}
%
%\authorrunning{} % abbreviated author list (for running head)
%
\institute{Department of Computer Science, \\
Aalto University School of Science, 
Espoo, Finland\\
\email{\{nam.vuhoang,seppo.torma\}@aalto.fi}}

@@

The volume of data produced in the construction industry is growing all the time. However, building data is currently difficult to access in an open manner. We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology. Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets  by using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components. Moreover, it paves the way to a completely distributed and loosely coupled architecture for building information management. Finally, opening selected parts of construction data to a wider set of stakeholders creates opportunities for new and innovative applications in the smart cities of the future.

\keywords{IFC, BIM, Ontology, Linked Data, OWL, RDF}

>>>>

\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}

%\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}

@@

% ifcOWL layer names
\newcommand{\simple}{\textbf{Simple}}
\newcommand{\standard}{\textbf{Standard}}
\newcommand{\extended}{\textbf{Extended}}

\newcommand{\ifcowl}{\textbf{ifcOWL}}
\newcommand{\ifcsimple}{\ifcowl{}-\simple{}}
\newcommand{\ifcstandard}{\ifcowl{}-\standard{}}
\newcommand{\ifcextended}{\ifcowl{}-\extended{}}

@@

%\vspace{1cm}
%
\title{Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology}
%
\titlerunning{Building ontology}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Nam Vu Hoang \and Seppo T\"orm\"a}
%
%\authorrunning{} % abbreviated author list (for running head)
%
\institute{Department of Computer Science, \\
Aalto University School of Science, 
Espoo, Finland\\
\email{\{nam.vuhoang,seppo.torma\}@aalto.fi}}

@@

The volume of data produced in the construction industry is rapidly growing but, as of now, it is not available for open and flexible access and utilization. We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology. Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets  by using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components. It paves the way to a distributed and loosely coupled architecture for building information management. Opening selected parts of construction data to a wider set of stakeholders creates opportunities for innovative applications in the smart cities of the future.

\keywords{Ontology, Linked Data, OWL, RDF, IFC, BIM}

### 2015-04-29 article-01-introduction.tex
Wrote sth about ifcRDF

Currently, a huge amount of basic design information about buildings are produced  
using modern Building Information Modeling (BIM) tools. There is a widely adopted standard, 
IFC (Industry Foundation Classes) that provides a common representation of building designs developed in different design disciplines: architecture, structural engineering,
MEP engineering, and so on. Nowadays, the IFC standard is supported by all major BIM
vendors; their BIM tools can export building models as IFC files that can be exchanged with other
parties.

The BIM models could provide a natural framework to organize and manage information gathered from other sources. However, a big obstacle to that is the difficulty to access and utilize BIM models. 
In the realm of IFC the sharing of building models is based on point-to-point file exchange which 
requires repeated manual work from designers. Information cannot be used in an online, granular, and interlinked manner, which provides a poor basis for open utilization of building information and the emergence of new, innovative applications.

To make building information more readily accessible, we have developed a \emph{WebIFC converter} to derive an OWL ontology from the IFC schema (\emph{ifcOWL}), and RDF datasets from IFC files (\emph{ifcRDF}). The ifcOWL ontology is used to classify concepts, characterize possible relationships, and define constraints in ifcRDF datasets which replicate the content of original IFC data files. All terms defined in the ifcOWL ontology must therefore be based on the corresponding constructs in the original IFC schema. The conversion from IFC data files to RDF is straightforward and can be done without information loss.

Ideally, there should also be just one ifcOWL ontology for each version of the IFC schema (Fig. \ref{fig:ifcOWL-layers}, a). However, the conversion from the EXPRESS data definition language (that is used to specify the IFC schema) to OWL can be done in several different ways for the following reasons. Firstly, there is a \emph{mismatch} between the modeling constructs between EXPRESS and OWL. Secondly, for most practical purposes, the IFC data can be considered \emph{valid and consistent} with respect to the schema, since the IFC models are exported from BIM tools with certified converters. Thirdly, as the master data of the design models will remain in the native formats of the BIM tools and evolves over time, the icfRDF can in practice only be used in a \emph{read-only} fashion. Fourthly, the conversion of all possible IFC schema information will lead into a \emph{extensive} OWL ontology that mostly contains information not needed in practical applications or use cases.

Consequently, we have converted the IFC schema into a \emph{flexible, multilayer ifcOWL ontology}. Different representational constructs are introduced at different levels. The ifcRDF is compatible with any of the layers; different layers just provide different amounts of type information that can be used with RDF.

\begin{figure}[h]
\centering
\includegraphics{images/ifcOWL-approaches.jpg}
\caption{Two approaches of creating and using ifcOWL ontology}
\label{fig:ifcOWL-layers}
\end{figure}

Different levels of details of the ifcOWL ontology are the following:

\begin{enumerate}
\item
    \textbf{ifcOWL-Lite} (compatible with OWL 2 EL, OWL 2 QL, OWL 2 RL)
    \begin{itemize}
        \item Schema meta data
        \item All type definitions and hierarchies
        \item All entity type properties (names and range types)
    \end{itemize}
\item
    \textbf{ifcOWL-Standard} (compatible with OWL 2 RL)
    \begin{itemize}
        \item Schema meta data
        \item All entity type keys
        \item All entity type inverse properties
    \end{itemize}
\item
    \textbf{ifcOWL-Advanced} (compatible with OWL 2 RL)
    \begin{itemize}
        \item All properties' cardinality constraints
        \item All list types' cardinality constraints
    \end{itemize}
\end{enumerate}

Since all ifcRDF datasets are compatible with all layers of the ifcOWL ontology (Fig. \ref{fig:ifcOWL-layers}, b), a user can choose the best one based on the requirements of the concrete use case and tools. For instance, many Linked Data applications that need to query entity properties, extract the type hierarchy, or convert ifcRDF datasets back to object-oriented data models it is sufficient to use the \textbf{ifcOWL-Lite} ontology. In case of reasoning about data that includes inverse properties, it is recommended to use \textbf{ifcOWL-Standard} ontology. If an RDF dataset needs to be modified and data validated using type and cardinality constraints, \textbf{ifcOWL-Advanced} ontology is the right choice.

>>>>

Currently, huge amounts of information about building designs is produced  
using Building Information Modeling (BIM) tools. A widely adopted standard, 
IFC (Industry Foundation Classes) provides a common representation for BIM models developed in 
different design disciplines. All major BIM tools support IFC and can export BIM models to IFC files for exchange with other parties.

BIM models could provide a natural framework to organize and manage information gathered from other 
sources. However, a big obstacle is the difficulty to access and utilize BIM models. 
In the realm of IFC the sharing of building models is based on point-to-point file exchange which 
requires repeated manual work from designers. Information cannot be used in an online, granular, and 
interlinked manner which provides a poor basis for open utilization of building information and the 
emergence of new, innovative applications.

To make building designs more openly accessible over the Web, we have developed the 
\emph{WebIFC converter} to 
derive OWL ontologies from the IFC schema (\ifcowl{}), and RDF datasets from IFC files 
(\ifcrdf{}). The ifcOWL ontology is used to classify concepts, characterize possible relationships, 
and define constraints in \ifcrdf{} datasets which replicate the content of original IFC data files. 
All terms defined in the \ifcowl{} ontology must therefore be based on the corresponding constructs in 
the original IFC schema. The conversion from IFC data files to RDF is straightforward and can be done 
without information loss.

Ideally, there should also be just one \ifcowl{} ontology for each version of the IFC schema. However,
there are possibilities and reasons to convert the IFC schema to OWL in several different ways for
the following reasons. Firstly, there is a \emph{mismatch} between the modeling constructs of OWL and 
EXPRESS data definition language used to specify the IFC schema, which creates
different conversion options. Secondly, a simple version of \ifcowl{} \emph{compatible with
restricted OWL profiles} (OWL 2 EL or QL) is useful to support ontology
reasoning or inexact reasoning. Thirdly, for most practical purposes only a subset of ontology
information is ever required, since no validation or checking of \ifcrdf{} data is needed; it can
be considered \emph{practically valid} with respect to the schema, since IFC files are exported
from BIM tools with certified converters, and icfRDF can sensibly only be used in a
\emph{read-only} fashion, since BIM models will be maintained in the native formats of BIM
tools. Fourthly, a full conversion of the IFC schema will make \ifcowl{}
\emph{unnecessarily large}, and difficult to understand or manipulate.

\begin{figure}[h]
\centering
\includegraphics[angle=0,width=1.0\textwidth]{images/ifcOWL-multilayers.png}
%\includegraphics{images/ifcOWL-multilayers.jpg}
\caption{Multiple layers of \ifcowl{} ontology}
\label{fig:ifcOWL-layers}
\end{figure}

Consequently, we have converted the IFC schema into a \emph{flexible, multilayer \ifcowl{} ontology}
in a way that \emph{\ifcrdf{} is compatible with any of the layers} (Fig. \ref{fig:ifcOWL-layers}). 
New representational constructs are introduced at more comprehensive levels which therefore 
provide increasing detailed type information about \ifcrdf{} data. Different levels of details of the \ifcowl{} ontology are the following:

\noindent
\begin{tabular}{|l|l|l|}
\hline
\textbf{Layer} & Profiles & Included schema information \\
\hline
\hline
\name{ifcOWL-Simple} & OWL 2 EL & Schema meta data \\
& OWL 2 QL &  All type definitions and hierarchies \\
& OWL 2 RL & All entity type properties (names, range types) \\
\hline 
\name{ifcOWL-Standard} & OWL 2 RL & Schema meta data \\
& & All entity type keys \\
& & All entity type inverse properties \\
\hline
\name{ifcOWL-Extended} & OWL 2 RL & All cardinality constraints of properties \\
& & All cardinality constraints of list types \\
\hline
\end{tabular}

Since all \ifcrdf{} datasets are compatible with all layers of the \ifcowl{} ontology 
(Fig. \ref{fig:ifcOWL-layers}), a user can choose the best one based on the requirements of 
the concrete use case and tools. For instance, many Linked Data applications that need to 
query entity properties, extract the type hierarchy, or convert \ifcrdf{} datasets back to 
object-oriented data models it is sufficient to use the \name{ifcOWL-Simple} ontology. 
In case of reasoning about data that includes inverse properties, it is recommended to use 
\name{ifcOWL-Standard} ontology. If an RDF dataset needs to be modified and data validated 
using type and cardinality constraints, \name{ifcOWL-Extended} ontology is the correct choice.

### 2015-04-29 article-02-main.tex
Wrote sth about ifcRDF

\section{Generating ifcOWL Ontologies From IFC Specifications}
\label{sec:ifcOWL}

\subsection{Common Conversion Rules}

\begin{ontologyRule}[OWL profile]
\ifclite{} layer must be compatible with any OWL 2 DL profile: OWL 2 EL, OWL 2 QL or OWL 2 RL. \ifcstandard{}, \ifcadvanced{} must be compatible with OWL 2 RL\footnote{Our WebIFC converter tool allows forced choice of other OWL profiles for any layer: OWL Lite, OWL DL, OWL Full, and OWL 2 Full}.
\end{ontologyRule}

\TODO{Explain: why an OWL 2 profile has to be being chosen, how that choice will impact to the reasoning and querying performance, whether we can use other OWL 2 EL or QL for ifcOWL-Lite ontology}

OWL 2 RL has been chosen for \ifcstandard{}, \ifcadvanced{} because in unlike OWL 2 EL and OWL 2 QL, it supports such needed things as: union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}), etc.\cite{w3c:owl2-profiles}.

\begin{ontologyRule}[Domain and range information]
Property domain and range constraints specified by \name{rdfs:domain} and \name{rdfs:range} should not be included in ifcOWL ontologies.
\end{ontologyRule}

Properties \name{rdfs:\-domain} and \name{rdfs:\-range} are mainly used to infer types, but not to check type constraints like in programming languages \cite{w3c:owl-guide}. Moreover, the basic facilities given by them do not provide any direct way to indicate property restrictions that are local to a class \cite{w3c:rdf-schema}. Although it is possible to combine use \name{rdfs:domain} and \name{rdfs:range} with sub-property hierarchies, direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.

\begin{ontologyRule}[Namespaces]
All constructs related to EXPRESS, STEP or IFC specifications must be belong to namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly. At this moment the namespaces are defined as below (istead of \name{IFCXXX} there should be IFC versions, e.g. \name{IFC2x3} or \name{IFC4}):

%\begin{lstlisting}[caption={Namespace definitions},label=lst:ifcOWL-namespaces]
\begin{lstlisting}
@prefix expr: <http://drumbeat.cs.hut.fi/owl/EXPRESS#>
@prefix step: <http://drumbeat.cs.hut.fi/owl/STEP#>
@prefix ifc: <http://drumbeat.cs.hut.fi/owl/IFCXXX#>
\end{lstlisting}
\end{ontologyRule}

@@

\begin{ontologyRule}[Simple data\-types]
All simple data\-types except \emph{BOOLEAN} and \emph{LOGICAL} are declared as the same as their most similar primitive XSD data\-types which are supported by OWL 2 \cite{w3c:owl2-profiles} and are among preferred standard data\-types in Linked Data Platform \cite{w3c:ldp-best-practices}:

% \begin{lstlisting}[caption={Simple data types}, label=lst:ifcOWL-simple-types]
\begin{lstlisting}
expr:STRING, ifc:IfcGlobalUniqueId owl:sameAs xsd:string .
expr:BINARY owl:sameAs xsd:hexBinary .
expr:INTEGER owl:sameAs xsd:integer .
expr:REAL, expr:NUMBER owl:sameAs xsd:double .
\end{lstlisting}
\end{ontologyRule}

@@

\begin{ontologyRule}[Simple data\-types with sizes, ifcOWL layers: all]If \name{STRING} and \name{BINARY} data\-types are specified with a size parameter, i.e. \name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then in layers \lite{} and \standard{} they are defined as equivalent to data\-types without size. But in layer \advanced{}, the size parameter must be declared by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2}, for instance:

% \begin{lstlisting}[caption={Size constraint of simple datatypes},label=lst:ifcOWL-simple-types-with-size]
\begin{lstlisting}
# only for Lite, Standard layers
expr:STRING22 owl:sameAs expr:STRING .

# only for Advanced layer
expr:STRING22 owl:equivalentClass
    [   a rdfs:Datatype ;
        owl:onDatatype  xsd:string ;
        owl:withRestrictions ( [ xsd:length "22"^^xsd:integer ] ) ] .
\end{lstlisting}
\end{ontologyRule}

@@

\begin{ontologyRule}[Enumeration data\-types]
Each enumeration data\-type in \textbf{ifc\-OWL} is defined as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}. All possible values of enumeration data\-types are defined as named individuals. Enumeration data\-type memberships are defined by means of \name{rdf:type} in the \lite{} layer, and by means of \name{owl:oneOf} in \standard{} and \advanced{} layers as shown below:
% \begin{lstlisting}[caption={Enumeration data\-types}, label=lst:ifcOWL-enum-types]
\begin{lstlisting}
# only for Lite layer
ifc:IfcAssemblyPlaceEnum a owl:Class ; rdfs:subclassOf expr:EnumClass .
ifc:SITE, ifc:FACTORY, ifc:NOTDEFINED rdf:type ifc:IfcAddressTypeEnum .

# only for Standard, Advanced layers
ifc:IfcAssemblyPlaceEnum a owl:Class ;  rdfs:subclassOf expr:EnumClass ;
    owl:oneOf ( ifc:SITE ifc:FACTORY ifc:NOTDEFINED ) .
\end{lstlisting}
\end{ontologyRule}

@@

\begin{ontologyRule}[Select data\-types]
A select data\-type must be converted to an \name{owl:Class} which is a subclass of \name{expr:\-Select\-Class}. Union-member data\-types are defined by means of \name{rdfs:\-subclassOf} in the \lite{} layer, and by means of \name{owl:unionOf} in \standard{} and \advanced{} layers as shown in the listing below.
\end{ontologyRule}

%\begin{lstlisting}[caption={Select data\-types}, label=lst:ifcOWL-select-types]
\begin{lstlisting}
# only for Lite layer
ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass .
ifc:IfcCartesianPoint, ifc:IfcParameterValue
    rdfs:subclassOf ifc:IfcTrimmingSelect .

# only for Standard, Advanced layers
ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass ;
    owl:unionOf ( ifc:IfcCartesianPoint ifc:IfcParameterValue ) ] .
\end{lstlisting}

@@

\begin{ontologyRule}[Defined data\-types]
If the defined data\-type is based on a simple data\-type then in ifcOWL it is declared as a \name{owl:Class} which is a subclass of \name{expr:\-Defined\-Class} has a single property \name{rdf:value} with that simple data\-type. If the defined data\-type is based on another defined data\-type then it is declared as a subclass of the \name{owl:Class} of that defined data\-type, for instance:

@@

ifc:IfcPositiveLengthMeasure a owl:Class ;
    rdfs:subClassOf ifc:IfcLengthMeasure .
\end{lstlisting}
\end{ontologyRule}

@@

\begin{ontologyRule}[Entity data\-types]An entity data\-type is converted to an \name{owl:\-Class} which is subclass of \name{expr:\-Entity\-Class} or its superclass. Entity attributes are defined by means of property \name{owl:on\-Pro\-per\-ty}. If an attribute is optional then its minimum qualified cardinality is 0, otherwise -- 1. Maximum qualified of all attributes are always 1. In case if an attribute is a list or a set of some data\-type then its type is considered as an external aggregated data\-type, for example:
%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
\begin{lstlisting}
ifc:IfcProperty a owl:Class ; rdfs:subClassOf expr:EntityClass .

ifc:IfcComplexProperty a owl:Class ; rdfs:subClassOf ifc:IfcProperty ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
                      owl:allValuesFrom ifc:LIST_1_3_IfcLengthMeasure ] ;
    # only for Advanced layer
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
                      owl:cardinality "1"^^xsd:integer ] .

# only for Standard & Advanced layers
[] a owl:AllDisjointClasses ;
    owl:members ( ifc:IfcComplexProperty ifc:IfcSimpleProperty ) .
\end{lstlisting}
\end{ontologyRule}

@@

# only for layer Advanced:
 ifc:SET_2_UNDEFINED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
    owl:onProperty expr:slot ; owl:minCardinality "2"^^xsd:integer ] .

@@

\section{Generating ifcRDF Datasets From IFC Data}

\begin{lstlisting}
:GUID_1916794F-5605-4499-9AB0-F155DE8D3B6C
    a               ifc:IfcPropertySet ; # entity type
    ifc:globalId    [ a ifc:IfcGloballyUniqueId ; # defined/select type
                     rdf:value "0P5dbFLWL4cPgmyLNUZJji"^^xsd:string ] ;
    ifc:hasProperties  _:b11 .

@@

\subsection{Metadata}

\TODO {Recommendations about model namespaces, naming individuals, defined types of select types }

>>>>

\section{Generating ifcOWL Ontologies From IFC Specifications}
\label{sec:ifcOWL}
This section describes the derivation of the ifcOWL ontologies from the IFC specification. Since there are different options for the conversion, the choices are presented as conversion principles with justiciations.

\subsection{Common Conversion Principles}

\begin{principle}[Simple OWL profiles]
\ifcowl{} should be compatible with as simple OWL 2 profiles as possible (to support a maximum variety of reasoners and other tools). 
\end{principle}

The first layer \ifcsimple{} can be specified to be compatible with any OWL 2 DL profile: OWL 2 EL, OWL 2 QL or OWL 2 RL. The other layers \ifcstandard{} and \ifcextended{} are compatible with OWL 2 RL\footnote{Our WebIFC converter tool allows forced choice of other OWL profiles for any layer: OWL Lite, OWL DL, OWL Full, and OWL 2 Full}. OWL 2 RL needed because it -- unlike OWL 2 EL or QL -- supports 
union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), and anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}) required by the IFC schema. \cite{w3c:owl2-profiles}

\begin{principle}[No domain and range restrictions]
Property names should be kept simple and therefore domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) should not be included in \ifcowl{} ontologies.
\end{principle}

In the IFC schema the property names are \emph{local} to entity classes while in RDF they are \emph{global}. 
In IFC the same property name can appear in multiple different classes. If the domain and range restrictions
were included in  \ifcowl{}, the property names would need to be qualified with entity class names. This 
would reduce the readability and usability of the \ifcowl{} significantly. In \ifcowl{} the domain and 
range restrictions would mainly be used to infer types, since there is no need to check type constraints 
like in programming languages \cite{w3c:owl-guide}. Moreover, the basic facilities given by them do not 
provide any direct way to indicate property restrictions that are local to a class \cite{w3c:rdf-schema}. 
Although it is possible to combine use \name{rdfs:domain} and \name{rdfs:range} with sub-property hierarchies, 
direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.

\begin{principle}[Namespaces]
All constructs related to EXPRESS, STEP or IFC specifications must be belong to namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g. \name{IFC4_ADD1}):
%\name{IFC2x3} or \name{IFC4}):
%At this moment the namespaces are defined as below (instead of \name{IFCXXX} there should be IFC versions, e.g. 
%\name{IFC2x3} or \name{IFC4}):

%\begin{lstlisting}[caption={Namespace definitions},label=lst:ifcOWL-namespaces]
\begin{lstlisting}
@prefix expr: <http://drumbeat.cs.hut.fi/owl/EXPRESS#>
@prefix step: <http://drumbeat.cs.hut.fi/owl/STEP#>
@prefix ifc: <http://drumbeat.cs.hut.fi/owl/IFCXXX#>
\end{lstlisting}
\end{principle}

@@

\begin{principle}[Simple data\-types]
All simple data\-types except \emph{BOOLEAN} and \emph{LOGICAL} are declared as the same as their most similar primitive XSD data\-types which are supported by OWL 2 \cite{w3c:owl2-profiles} and are among preferred standard data\-types in Linked Data Platform \cite{w3c:ldp-best-practices}:

% \begin{lstlisting}[caption={Simple data types}, label=lst:ifcOWL-simple-types]
\begin{lstlisting}
expr:STRING, ifc:IfcGlobalUniqueId owl:sameAs xsd:string .
expr:BINARY owl:sameAs xsd:hexBinary .
expr:INTEGER owl:sameAs xsd:integer .
expr:REAL, expr:NUMBER owl:sameAs xsd:double .
\end{lstlisting}
\end{principle}

@@

\begin{principle}[Simple data\-types with sizes, ifcOWL layers: all]If \name{STRING} and \name{BINARY} data\-types are specified with a size parameter, i.e. \name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then in layers \simple{} and \standard{} they are defined as equivalent to data\-types without size. But in layer \extended{}, the size parameter must be declared by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2}, for instance:

% \begin{lstlisting}[caption={Size constraint of simple datatypes},label=lst:ifcOWL-simple-types-with-size]
\begin{lstlisting}
# only for Simple, Standard layers
expr:STRING22 owl:sameAs expr:STRING .

# only for Extended layer
expr:STRING22 owl:equivalentClass
    [   a rdfs:Datatype ;
        owl:onDatatype  xsd:string ;
        owl:withRestrictions ( [ xsd:length "22"^^xsd:integer ] ) ] .
\end{lstlisting}
\end{principle}

@@

\begin{principle}[Enumeration data\-types]
Each enumeration data\-type in \textbf{ifc\-OWL} is defined as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}. All possible values of enumeration data\-types are defined as named individuals. Enumeration data\-type memberships are defined by means of \name{rdf:type} in the \simple{} layer, and by means of \name{owl:oneOf} in \standard{} and \extended{} layers as shown below:
% \begin{lstlisting}[caption={Enumeration data\-types}, label=lst:ifcOWL-enum-types]
\begin{lstlisting}
# only for Simple layer
ifc:IfcAssemblyPlaceEnum a owl:Class ; rdfs:subclassOf expr:EnumClass .
ifc:SITE, ifc:FACTORY, ifc:NOTDEFINED rdf:type ifc:IfcAddressTypeEnum .

# only for Standard, Extended layers
ifc:IfcAssemblyPlaceEnum a owl:Class ;  rdfs:subclassOf expr:EnumClass ;
    owl:oneOf ( ifc:SITE ifc:FACTORY ifc:NOTDEFINED ) .
\end{lstlisting}
\end{principle}

@@

\begin{principle}[Select data\-types]
A select data\-type must be converted to an \name{owl:Class} which is a subclass of \name{expr:\-Select\-Class}. Union-member data\-types are defined by means of \name{rdfs:\-subclassOf} in the \simple{} layer, and by means of \name{owl:unionOf} in \standard{} and \extended{} layers as shown in the listing below.
\end{principle}

%\begin{lstlisting}[caption={Select data\-types}, label=lst:ifcOWL-select-types]
\begin{lstlisting}
# only for Simple layer
ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass .
ifc:IfcCartesianPoint, ifc:IfcParameterValue
    rdfs:subclassOf ifc:IfcTrimmingSelect .

# only for Standard, Extended layers
ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass ;
    owl:unionOf ( ifc:IfcCartesianPoint ifc:IfcParameterValue ) ] .
\end{lstlisting}

@@

\begin{principle}[Defined data\-types]
If the defined data\-type is based on a simple data\-type then in ifcOWL it is declared as a \name{owl:Class} which is a subclass of \name{expr:\-Defined\-Class} has a single property \name{rdf:value} with that simple data\-type. If the defined data\-type is based on another defined data\-type then it is declared as a subclass of the \name{owl:Class} of that defined data\-type, for instance:

@@

ifc:IfcPositiveLengthMeasure a owl:Class ;
    rdfs:subClassOf ifc:IfcLengthMeasure .
\end{lstlisting}
\end{principle}

@@

\begin{principle}[Entity data\-types]An entity data\-type is converted to an \name{owl:\-Class} which is subclass of \name{expr:\-Entity\-Class} or its superclass. Entity attributes are defined by means of property \name{owl:on\-Pro\-per\-ty}. If an attribute is optional then its minimum qualified cardinality is 0, otherwise -- 1. Maximum qualified of all attributes are always 1. In case if an attribute is a list or a set of some data\-type then its type is considered as an external aggregated data\-type, for example:
%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
\begin{lstlisting}
ifc:IfcProperty a owl:Class ; rdfs:subClassOf expr:EntityClass .

ifc:IfcComplexProperty a owl:Class ; rdfs:subClassOf ifc:IfcProperty ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
                      owl:allValuesFrom ifc:LIST_1_3_IfcLengthMeasure ] ;
    # only for Extended layer
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
                      owl:cardinality "1"^^xsd:integer ] .

# only for Standard & Extended layers
[] a owl:AllDisjointClasses ;
    owl:members ( ifc:IfcComplexProperty ifc:IfcSimpleProperty ) .
\end{lstlisting}
\end{principle}

@@

# only for layer Extended:
 ifc:SET_2_UNDEFINED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
    owl:onProperty expr:slot ; owl:minCardinality "2"^^xsd:integer ] .

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Generating ifcRDF Datasets From IFC Data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Generating ifcRDF Datasets From IFC Data}
\label{sec:ifcRDF}

\subsection{Dataset Metadata}
\label{subsec:ifcRDF-metadata}

IFC data are exchanged by using ``STEP physical files'' (SPF). SPF format is defined in ISO 10303-21 \cite{wiki:step-file}. A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription} \name{FileName} and \name{FileSchema} described in \cite{buildingSMART:ifc-header}. They are exported as objects of STEP entities, some attributes are duplicated as Dublin Core (DC) metadata annotations, for example:

\begin{lstlisting}
<http://example.org>  a void:DataSet ;
    dcterms:created "2015-02-19T13:22:15Z"^^xsd:dateTime ;
    dcterms:creator "John Smith"^^xsd:string .
    step:fileName  [ a step:FileName ;
        step:author "John Smith"^^xsd:string ;
        step:originating_system "Tekla Structures 20.1"^^xsd:string ; ] ;
        step:time_stamp "2015-02-19T13:22:15Z"^^xsd:dateTime  ] .
\end{lstlisting}

In a SPF data section each entity is declared with a line number, an entity data\-type and a list of values. All entities of \name{Ifc\-Root}-derived types are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} must be an normal GUID, not a compressed IFC GUID which may include a sign \name{\$}. Regarding to other entities, user can choose whether to export them as anonymous individuals or name them in a format like \name{LINE\_\textless{}lineNumber\textgreater{}}. Note that line numbers are used only for linking entities inside a SPF, and they are unstable. Naming is required when the dataset is used together with OWL 2 EL or OWL 2 QL because these profiles do not allow anonymous individuals. The namespace of named entities is also defined by user. Below is an example of IFC data exported to an RDF dataset which is compatible with OWL 2 RL:

\begin{lstlisting}
:GUID_1916794F-5605-4499-9AB0-F155DE8D3B6C
    a               ifc:IfcPropertySet ; # entity type
    ifc:globalId    [ a ifc:IfcGloballyUniqueId ; # defined type
                     rdf:value "0P5dbFLWL4cPgmyLNUZJji"^^xsd:string ] ;
    ifc:hasProperties  _:b11 .

@@

%Data\-type of an attribute value is declared obviously only if the attribute has a select data\-type and the value has a defined data\-type.

\section{Conclusion}
\label{sec:conclusion}
